---
title: "Estimation of Behavioral Entropy Rate : R-Software"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# $\texttt{ccber}$: an `R` Package for the Estimation of Behavioral Entropy Rate - Developed for the Conte Center @ UCI

See reference: Davis, E.P., Stout, S.A., Molet, J., Vegetabile, B., Glynn, L.M., Sandman, C.A., Heins, K., Stern, H., Baram, T.Z. (in press). **Exposure to unpredictable maternal sensory signals influences cognitive development across-species**. _Proceedings of the National Academy of Sciences_.

## Installing `ccber`

The package `devtools` is required to install this `R` package from this Github repository.  Install this package first if it is not already installed.

```{r, echo=TRUE, eval=FALSE}
install.packages('devtools', dependencies = TRUE)
```

Once that package has been installed, use the following to install `ccber`

```{r, echo=TRUE, eval=FALSE}
devtools::install_github('bvegetabile/ccber')
```

Load the package to begin analysis!

```{r, echo=TRUE, eval=TRUE}
library('ccber')
```

## Using `ccber` with Observer Files

To use `ccber` to estimate behavioral entropy rate, see the Conte Center website for a description of how to set up files and record observations.  The input files for these functions are described there...

### Running a single file

To run `ccber` with a single file use the following function
```{r, echo=TRUE, eval=FALSE}
ber_analyze_file(f_loc,
                 plot_all=F,
                 plots_to_file=F,
                 tactile_padding = 1.0,
                 auditory_padding = 1.0,
                 behavior_types=list("mom_auditory_types" = c('Vocal'),
                                     "mom_tactile_types" = c('TouchBaby',
                                                            'HoldingBaby'),
                                     "mom_visual_types" = c('ManipulatingObject'),
                                     "baby_visual_types" = c('LookAtMomActivity'),
                                     "missing_types" = c('CantTellHolding',
                                                         'ActivityNotVisible',
                                                         'CantTellLooking')),
                 missing_threshold = 0.1)
```
The variables are described below:

* `f_loc` : Location of the file to be analyzed
* `plot_all` : Logical variable to plot diagnostic plots for this individual. Defaults to `FALSE`
* `plots_to_file` : Logical variable which plots diagnostic plots to a file.  Currently unused.
* `tactile_padding` : Padding to be applied to the "event" types of the Observer software.  Padding is right-adjusted and Defaults to 1 second.
* `auditory_padding` : Padding to be applied to the "event" types of the Observer software.  Padding is right-adjusted and Defaults to 1 second.
* `behavior_types` : List specifying the "auditory", "tactile", and "visual" behaviors to subset on.  The "missing_types" are description of tags that define missing data.  
* `missing_threshold` : Proportion value that defines what the threshold of missingness is to include the file or not.  Set to 0.1 reflecting 10% of missingness is acceptable based upon the tags in `behavior_types`.

### Running on a Directory of Files

To expedite processing of many files an additionally function is provided to analyze an entire directory of Excel files.  The function `ber_analyze_dir` is similar to the previous function and takes as input `dir_loc`.

```{r, echo=T, eval=F}
ber_analyze_dir(dir_loc,
                tactile_padding = 1.0,
                auditory_padding = 1.0,
                behavior_types=list("mom_auditory_types" = c('Vocal'),
                                    "mom_tactile_types" = c('TouchBaby',
                                                            'HoldingBaby'),
                                    "mom_visual_types" = c('ManipulatingObject'),
                                    "baby_visual_types" = c('LookAtMomActivity'),
                                    "missing_types" = c('CantTellHolding',
                                                        'ActivityNotVisible',
                                                        'CantTellLooking')),
                missing_threshold = 0.1)
```

The inputs that are described in the previous section are the same and are passed as input to multiple calls of `ber_analyze_file`.  

## An Example of How to Estimate Entropy Rate using `ccber`

Consider the following transition matrix of a first-order Markov chain with three states,

```{r, echo=TRUE, eval=TRUE}
P = matrix(c(0.2, 0.3, 0.5, 
             0.7, 0.1, 0.2,
             0.2, 0.2, 0.6), 3,3, byrow = T)
```

We can simulate from a Markov process with this using the function `SimulateMarkovChain`

```{r, echo=T, eval=T}
mc_chain <- SimulateMarkovChain(trans_mat = P, n_sims = 5000)
head(mc_chain, n = 20)
```

From this we can calculate a matrix of transition counts 

```{r, echo=T, eval=T}
tc <- CalcTransitionCounts(mc_chain)
tc
```

And then estimate a transition matrix,

```{r, echo=T, eval=T}
tm <- CalcTransitionMatrix(tc)
tm
```

which agrees fairly well with the true `P`.  Additionally we can estimate the stationary distribution of the process in a one of two ways.  The first way is an empirical estimate from the observed sequence.  

```{r, echo=T, eval=T}
emp_sm <- CalcEmpiricalStationary(mc_chain, state_space = 1:3)
emp_sm
```

The second way is an eigendecomposition of the observed transition matrix, though the preferred method is through the empirical estimation procedure.

```{r, echo=T, eval=T}
eig_sm <- CalcEigenStationary(tm)
eig_sm
```

Using both the stationary distribution estimate and the estimate of the transition matrix, the entropy rate of the process can be estimated using the following commands

```{r, echo=T, eval=T}
entrate1 <- CalcMarkovEntropyRate(tm, emp_sm)
entrate1
```

```{r, echo=T, eval=T}
entrate2 <- CalcMarkovEntropyRate(tm, eig_sm)
entrate2
```

Both of these values agree very well with the true entropy rate,
```{r, echo=T, eval=T}
true_entropy_rate <- CalcMarkovEntropyRate(P, CalcEigenStationary(P))
true_entropy_rate
```


## There's a Quicker Way Than That...

If the method of estimation for the stationary distrbution is known, a more simple function is provided to estimate the entropy rate as well.

Both of these values agree very well with the true entropy rate,
```{r, echo=T, eval=T}
quicker_estimate <- CalcEntropyRate(mc_chain, 
                                    state_space = 1:3, 
                                    stat_method = "Empirical")
quicker_estimate
```

