---
title: "Estimation of Behavioral Entropy Rate : R-Software"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# $\texttt{ccber}$: an `R` Package for the Estimation of Behavioral Entropy Rate - Developed for the Conte Center @ UCI

See reference: _Early Life Exposure to Unpredictable Maternal Sensory Signals Influences Cognitive Development: A Cross-Species Approach_ 

## Installing `ccber`

The package `devtools` is required to install this `R` package from this Github repository.  Install this package first if it is not already installed.

```{r, echo=TRUE, eval=FALSE}
install.packages('devtools', dependencies = TRUE)
```

Once that package has been installed, use the following to install `ccber`

```{r, echo=TRUE, eval=FALSE}
devtools::install_github('bvegetabile/ccber')
```

Load the package to begin analysis!

```{r, echo=TRUE, eval=FALSE}
library('ccber')
```




## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
