---
title: "Estimation of Behavioral Entropy Rate : R-Software"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# $\texttt{ccber}$: an `R` Package for the Estimation of Behavioral Entropy Rate - Developed for the Conte Center @ UCI

See reference: _Early Life Exposure to Unpredictable Maternal Sensory Signals Influences Cognitive Development: A Cross-Species Approach_ Elysia Davis, Stephanie Stour, Jenny Molet, Brian Vegetabile, Laura Glynn, Curt Sandman, Kevin Heins, Hal Stern, and Tallie Baram: To appear in the Proceedings of the National Academy of Sciences

## Installing `ccber`

The package `devtools` is required to install this `R` package from this Github repository.  Install this package first if it is not already installed.

```{r, echo=TRUE, eval=FALSE}
install.packages('devtools', dependencies = TRUE)
```

Once that package has been installed, use the following to install `ccber`

```{r, echo=TRUE, eval=FALSE}
devtools::install_github('bvegetabile/ccber')
```

Load the package to begin analysis!

```{r, echo=TRUE, eval=TRUE}
library('ccber')
```

## Using `ccber` with Observer Files



## Example of Estimating Entropy Rate using `ccber`

Consider the following transition matrix of a first-order Markov chain with three states,

```{r, echo=TRUE, eval=TRUE}
P = matrix(c(0.2, 0.3, 0.5, 
             0.7, 0.1, 0.2,
             0.2, 0.2, 0.6), 3,3, byrow = T)
```

We can simulate from a Markov process with this using the function `SimulateMarkovChain`

```{r, echo=T, eval=T}
mc_chain <- SimulateMarkovChain(trans_mat = P, n_sims = 1000)
print(head(mc_chain, n = 20))
```
